A new simplified manual tour, with examples in mathematica

### 2022-04-01 paper/paper.Rmd
colour of selected variable

<!-- Purpose -->
Manual controls are especially useful for assessing sensitivity of structure to particular elements of the projection. There are many places where it is useful. In exploratory data analysis, where one sees clusters in a projection, can some variables be removed from the projection without affecting the clustering. For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance. These controls can also be used to interactively generate facetted plots [@XXX], or spatiotemporal glyphmaps [@XXX].

@@

@@

# Create manual tour example
p <- 4
d <- 2
set.seed(24)
A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
colnames(A) <- c("P1", "P2")
rownames(A) <- paste0("V", 1:p)
A <- tourr::orthonormalise(A)
plot_basis(A)

# Now set a new position
vchange <- 3
eps <- 0.02

# Iterate and save images
for (i in 1:11) {
  flname <- paste0("mtour_plots/mtour", i, ".png")
  A[vchange,] <- A[vchange,]*(1-i*eps)
  A <- tourr::orthonormalise(A)
  #png(flname, width=250, height=250)
  plot_basis(A)
  #dev.off()
}
```

>>>>

<!-- Purpose -->
Manual controls are especially useful for assessing sensitivity of structure to particular elements of the projection. There are many places where it is useful. In exploratory data analysis, where one sees clusters in a projection, can some variables be removed from the projection without affecting the clustering. For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance. These controls can also be used to interactively generate facetted plots [@XXX], or spatiotemporal glyphmaps [@XXX]. Having the user interact with a projection is extremely valuable for understanding high-dimensional data. However, these algorithms have two problems: (1) the pre-processing of creating a manipulation space overly complicates the algorithm, (2) extending to higher dimensional control is difficult.

Through experiments with the relatively new interactive graphics capabilities in mathematica(?), we have realized that there is a simpler approach, which is more direct, and extensible for generating user interaction. This paper explains this, and is organized as follows. The next section describes the new algorithm for manual control. This is followed by details on implementation. The applications section illustrate how these can be used.

@@

Describe the manip space, out of manual tour

1. Provide a k-D projection
2. Select variable to control
3. Detect or provide direction of change
4. Update projection, and orthonormalise

Need to think about how the values for selected variable are made exact, that is constrained orthonormalisation.

Need to think about checks, and error catching, maybe in implementation section. If selected variable has 0 coefficient, will that generate an orthonormalisation error?

@@

# Create manual tour example
p <- 4
d <- 2
set.seed(24)
A <- matrix(runif(p*d, min=-1), ncol=d, byrow=TRUE)
colnames(A) <- c("P1", "P2")
rownames(A) <- paste0("V", 1:p)
A <- tourr::orthonormalise(A)

# Now set a new position
vchange <- 3
eps <- 0.02
plot_basis(A, vchange)

# Iterate and save images
for (i in 1:11) {
  flname <- paste0("mtour_plots/mtour", i, ".png")
  A[vchange,] <- A[vchange,]*(1-i*eps)
  A <- tourr::orthonormalise(A)
  #png(flname, width=250, height=250)
  plot_basis(A, vchange)
  #dev.off()
}
```

### 2022-04-01 paper/paper.tex
colour of selected variable

Manual controls are especially useful for assessing sensitivity of
structure to particular elements of the projection. There are many
places where it is useful. In exploratory data analysis, where one sees
clusters in a projection, can some variables be removed from the
projection without affecting the clustering. For interpreting models,
one can reduce or increase a variable's contribution to examine the
variable importance. These controls can also be used to interactively
generate facetted plots \citep{XXX}, or spatiotemporal glyphmaps
\citep{XXX}.

@@

>>>>

Manual controls are especially useful for assessing sensitivity of
structure to particular elements of the projection. There are many
places where it is useful. In exploratory data analysis, where one sees
clusters in a projection, can some variables be removed from the
projection without affecting the clustering. For interpreting models,
one can reduce or increase a variable's contribution to examine the
variable importance. These controls can also be used to interactively
generate facetted plots \citep{XXX}, or spatiotemporal glyphmaps
\citep{XXX}. Having the user interact with a projection is extremely
valuable for understanding high-dimensional data. However, these
algorithms have two problems: (1) the pre-processing of creating a
manipulation space overly complicates the algorithm, (2) extending to
higher dimensional control is difficult.

Through experiments with the relatively new interactive graphics
capabilities in mathematica(?), we have realized that there is a simpler
approach, which is more direct, and extensible for generating user
interaction. This paper explains this, and is organized as follows. The
next section describes the new algorithm for manual control. This is
followed by details on implementation. The applications section
illustrate how these can be used.

@@

Describe the manip space, out of manual tour

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provide a k-D projection
\item
  Select variable to control
\item
  Detect or provide direction of change
\item
  Update projection, and orthonormalise
\end{enumerate}

Need to think about how the values for selected variable are made exact,
that is constrained orthonormalisation.

Need to think about checks, and error catching, maybe in implementation
section. If selected variable has 0 coefficient, will that generate an
orthonormalisation error?
